The Inner Mass-Ratio Distribution of Intermediate-Mass Stars

### 2016-02-25 paper/ms.tex
Updated abstract and intro as per Adams suggestions.

\begin{abstract}
Binary stars and higher-order multiple systems are a ubiquitous outcome of star formation, especially as the system mass increases. The distribution of mass ratios is a unique probe into the conditions of the collapsing molecular core as well as the circumstellar disks of the binary fragments. Inside $a \sim 1000$ AU the disks from the two forming stars can interact, and additionally companions can form directly through disk fragmentation. We should therefore expect the mass-ratio distribution of close companions to differ from that of wide companions. Unfortunately, close companions to A- and B-type stars are very difficult to detect and characterize because they are so much dimmer than the primary star. We present the results of a survey searching for companions to A- and B-type stars using the direct spectral detection technique, which is sensitive to late-type companions within $\sim 1''$ of the primary. We estimate the temperatures and masses of the 57 companion detections, 25 of which are new detections. We present the mass-ratio distribution for our sample, and find that it is has a robust maximum near mass ratios $q \sim 0.3$. This is a very different form for the mass ratio distribution, which is usually well-described by a power law, and indicates that close companions to intermediate-mass stars experience significantly different accretion histories or formation mechanisms than wide companions.
\end{abstract}
\keywords{binaries: spectroscopic, stars:early-type, stars:formation, stars:statistics}

@@

The overall multiplicity rate and the distributions of mass ratio, period, and eccentricity of a binary star population place important constraints on the mode of binary star formation. While the period and eccentricity are altered by dynamical processing in the birth cluster, the present-day mass ratio of a binary system is a tracer of its formation \citep{Parker2013}. Most binary stars are thought to form via core fragmentation \citep{Boss1979, Boss1986, Bate1995}, in which a collapsing core fragments into two or more individual protostars. The number and initial masses of the fragments are set by the total core mass, as well as its rotation, turbulence, and its temperature and density structure. If the fragments are well separated ($a \gtrsim 1000$ AU), they will evolve relatively independently of each other, accreting mass from the core material onto their own protostellar disks and then onto the protostars themselves. However close fragments ($a \sim 100$ AU) will interact with each other; the protostellar disk may be truncated, destabilized, or form into a circumbinary disk if the separation is small enough \citep{Bate1997}. In addition, an unstable disk can fragment to form low-mass companions \citep{Kratter2006, Stamatellos2011}. The mass ratios of close companions formed via either mechanism should be affected by preferential accretion. Most work has suggested that the disk material will preferentially accrete onto the lower mass companion \citep{Bate1997, BBB2002}; however, recent work has indicated that magnetic disk braking may result in preferential accretion onto the more massive component \citep{Zhao2013} instead. In either case, we would expect to find a mass ratio distribution for companions inside a few $100$ AU that differs from that of companions on wider orbits.

The mass ratio, period, and eccentricity distributions are fairly well-known for solar type stars \citep{Duquennoy1991, Raghavan2010} and cooler stars \citep{Fischer1992, Delfosse2004}. Interestingly, the mass ratio distribution appears to be invariant to separation for these stars \citep{Meyer2013}, contrary to the theoretical expectations. All of the distributions are much less certain for more massive stars. The reason for this is two-fold: first, more massive stars tend to more rare and farther away, meaning many of the companions are angularly close to the very bright primaries and difficult to detect with imaging techniques. Second, the primary stars tend to be rapid rotators, which limits radial velocity precision to $\sim 1 \mathrm{km\ s}^{-1}$. Worse, radial velocity monitoring can only measure a mass ratio if spectral lines from both components are visible; this typically suffers from the same flux ratio difficulty as imaging techniques.

Nonetheless, \citet{DeRosa2014} performed an adaptive optics imaging survey of nearby A-type stars, and found that the mass ratio distribution is well-described by a power law with large slope, indicating a very strong preference for low-mass companions. They also found initial evidence that the mass ratio distribution for companions inside $125$ AU has a much shallower power law slope than that of wide companions, and is consistent with flat. Unfortunately the close companion subsample contained only 18 binary systems, and the result is complicated by the inherent difficulty of detecting close companions with low mass ratios in an imaging survey.

Spectroscopic companions to chemically peculiar Am stars appear to have a mass ratio distribution which peaks near $q \sim 0.5$ \citep{Vuissoz2004}. Am stars are typically associated with binary companions, and are slow rotators due to tidal braking; they thus form a highly biased sample of intermediate-mass stars. It is however interesting that the inner mass ratio distribution around these chemically peculiar stars has an entirely different form than the distribution around chemically normal stars at wide separations probed by imaging.

>>>>

\begin{abstract}
Binary stars and higher-order multiple systems are a ubiquitous outcome of star formation, especially as the system mass increases. The distribution of mass ratios is a unique probe into the conditions of the collapsing cloud core and circumstellar disk(s) of the binary fragments. Inside $a \sim 1000$ AU the disks from the two forming stars can interact, and additionally companions can form directly through disk fragmentation. We should therefore expect the mass-ratio distribution of close companions to differ from that of wide companions. This prediction is difficult to test with intermediate-mass to massive primary stars because close companions are so much dimmer. We present the results of a survey searching for companions to A- and B-type stars using the direct spectral detection technique, which is sensitive to late-type companions within $\sim 1''$ of the primary and which has no inner working angle. We estimate the temperatures and masses of the 57 companions , 25 of which are new detections. We find that the mass-ratio distribution for our sample has a robust maximum near $q \sim 0.3$. Our mass-ratio distribution has a very different form than in previous work, where it is usually well-described by a power law, and indicates that close companions to intermediate-mass stars experience significantly different accretion histories or formation mechanisms than wide companions.
\end{abstract}
\keywords{binaries: spectroscopic, stars:early-type, stars:formation, stars:statistics}

@@

The overall multiplicity rate and the distributions of mass ratio, period, and eccentricity of a binary star population place important constraints on the mode of binary star formation. While the period and eccentricity are altered by dynamical processing in the birth cluster, the present-day mass ratio of a binary system is a direct result of its formation \citep{Parker2013}. Most binary stars are thought to form via core fragmentation \citep{Boss1979, Boss1986, Bate1995}, in which a collapsing core fragments into two or more individual protostars. The number and initial masses of the fragments are set by the total core mass, as well as its rotation, turbulence, and its temperature and density structure. If the fragments are well separated ($a \gtrsim 1000$ AU), they will evolve independently of each other, accreting mass from the core material onto their own protostellar disks and then onto the protostars themselves. However close fragments ($a \sim 100$ AU) will interact with each other; the protostellar disk may be truncated, destabilized, or form into a circumbinary disk if the separation is small enough \citep{Bate1997}. In addition, an unstable disk can fragment to form low-mass companions \citep{Kratter2006, Stamatellos2011}. The mass ratios of close companions formed via either mechanism should be affected by preferential accretion. Most work has suggested that the disk material will preferentially accrete onto the lower mass companion \citep{Bate1997, BBB2002}; however, recent work has indicated that magnetic disk braking may result in preferential accretion onto the more massive component \citep{Zhao2013} instead. In either case, we would expect to find a mass ratio distribution for companions inside a few $100$ AU that differs from that of companions on wider orbits.

The mass ratio, period, and eccentricity distributions are fairly well-known for solar type stars \citep{Duquennoy1991, Raghavan2010} and cooler stars \citep{Fischer1992, Delfosse2004}. Interestingly, the mass ratio distribution appears to be invariant to separation for these stars \citep{Meyer2013}, contrary to the theoretical expectations. All of the distributions are much less certain for more massive stars. The reason for this is two-fold: first, more massive stars tend to more rare and farther away, meaning many of the companions are angularly close to the very bright primaries and difficult to detect with imaging techniques. Second, the primary stars tend to be rapid rotators, which limits radial velocity precision to $\sim 1 \mathrm{km\ s}^{-1}$. Additionally, radial velocity monitoring can only measure a mass ratio if spectral lines from both components are visible; this typically suffers from the same flux ratio difficulty as imaging techniques.

Nonetheless, \citet{DeRosa2014} performed an adaptive optics imaging survey of nearby A-type stars, and found that the mass ratio distribution is well-described by a power law with large slope, indicating a very strong preference for low-mass companions. They also found initial evidence that the mass ratio distribution for companions inside $125$ AU has a much shallower power law slope than that of wide companions, and is consistent with flat. Their close companion subsample contained only 18 binary systems, and the result is complicated by the inherent difficulty of detecting close companions with low mass ratios in an imaging survey.

Radial velocity monitoring surveys can detect much closer companions than imaging surveys, but are typically only complete to low-mass companions if the primary is a slow rotator. Chemically peculiar Am stars are typically associated with binary companions, and are slow rotators due to tidal braking; they thus form a highly biased sample of intermediate-mass stars. Nonetheless, it is interesting to note that they have a mass-ratio distribution which peaks near $q \sim 0.5$ \citep{Vuissoz2004}, an entirely different form than the distribution found around chemically normal stars at wide separations.
